---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vsgseqtools

<!-- badges: start -->
<!-- badges: end -->

A collection of functions that helps analyze and tidy up the output from VSG sequencing. The purpose of this package is to be used specifically for analyzing data produced from the [VSGSeqPipeline](https://github.com/mugnierlab/VSGSeqPipeline). 

## Installation

``` r
library(devtools)
devtools::install_github("ABeav/vsgseqtools")
```

## Examples

There are several helpful tools in this package. Below is an example of a few tools that help read in and analyze VSGSeqPipeline results. 

```{r read, warning=FALSE, message=FALSE}
library(tidyverse) # The tidyverse is required
library(vsgseqtools)

## This function accurately reads in the results from the raw VSGSeqPipeline output
results <- vsg_read("mock_data/mock_RESULTS.txt")
head(results)
```

The `vsg_rename()` function allows you to add a column for the clusters produced by cd-hit in the VSGSeqpipeline. However, you do need to provide a reference that has each VSG name and its associated cluster name (I do this using this script [MatchVSGs.py](additional_resources/)).

```{r rename, warning=FALSE, message=FALSE}
vsg_rename(results, "mock_data/cluster_reference_table.txt") %>%
  select(mouse, day, tissue, Percent, cluster, VSG)
```

Often when analyzing VSG expression, we count the total number of VSGs in a sample. The `vsg_count` function helps streamline this with one line of code. 

```{r count, warning=FALSE, message=FALSE}
results %>%
  vsg_count(mouse, day, tissue) 

# This can also be plotted directly after

results %>%
  vsg_count(mouse, day, tissue) %>%
  ggplot(aes(x = tissue, y = vsg_count)) + 
  facet_wrap("day") +
  geom_boxplot() +
  geom_dotplot(binaxis='y', stackdir= "center" , dotsize=.6, position=position_dodge(.75)) +
  xlab("") +
  ylab("Number of VSGs") +
  theme(axis.text = element_text(angle = 30,hjust = 1, size = 16, color = 'black'))
```

It can be helpful to look at the most highly expressed VSG in each sample as well. I created a function, `vsg_max`, to help pull the maximum expressed VSG out easily. This is particularly useful to find the starting VSG expressed in an infection on day 6. 

```{r}
results %>%
  filter(day == "D6", tissue == "blood") %>%
  vsg_max(mouse) %>% 
  select(mouse, day, tissue, Percent, VSG)
```

The last function I wanted to highlight is little more niche. This function is helpful if you want to create a data frame that has includes all VSGs found and represents them in each sample. VSGs that were not found in a sample are represented by having a Percent of 0, meaning 0% of parasites expressed that VSG. 

```{r}
 vsg_expand(results, "samplename", "VSG") %>%
  select(samplename, Percent, VSG)
```

